GRN inference in multiple species

### 2015-05-04 paper1.tex
hi
abstract

\title{GRN inference in multiple species}
\author{Kari Y. Lam, Zachary M. Westrick, Frank Underwood}
\maketitle

\begin{abstract}
sup sup sup
\end{abstract}

@@

Where the second sum is over pairs of interactions with fusion constraints. Note that each term in the quantity being minimized is an L2 norm of a linear combination of gene expressions, so the entire problem can be solved through linear regression with a suitable design matrix [DETAILS SHOULD GO SOMEWHERE ELSE, OR NOWHERE]. The regulators of a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.

@@

\begin{equation}
\begin{array}{l}
\text{for each fused  } B^{S_1}_{g,k} \approx B^{S_2}_{h,l}  \text{ we define: } 
\\
\theta(0)=0
\\
\theta(t) = \vert B^{S_1}_{g,k} - B^{S_2}_{h,l} \vert
\\
\text{and introduce a fusion constraint } \lambda = \frac{p'(\theta(t))}{2\theta(t)} 
\\
\end{array}
\end{equation}

@@

\subsection{Beta scaling}
[INSERT JUSTIFICATION]
\\
In previous work, betas were rescaled as to form a matrix of confidence scores $S$ as follows
\begin{equation}
S_{i,j} = \frac{\sigma^2_{\text{full model for }y_j}}{\sigma^2_{\text{full model for }y_j \text{ without predictor }i}}
\end{equation}
Computing residuals with respect to the data alone would disregard information gained through fusion. Instead, we used an approximation
\begin{equation}
S_{i,j} = \frac{\sigma^2_{\text{full model for }y_j}}{\sigma^2_{\text{full model for }y_j} + B_{i,j}^2 \times var(TF_j)}
\end{equation}
When the residual is uncorrelated with the prediction, then removing regressor $i$ with weight $B_{i,j}^2$ from the model of $j$ will increase the residual by $B_{i,j}^2$ times the variance of regressor $i$. In the case of regularized regression, the regressor and residual need not be uncorrelated, so the approximation will not hold exactly. This scaling is similar to rescaling according to variance explained relative to an augmented design matrix that includes fusion constraints. We use an approximation because it is simpler to compute. The intuition is that a large $B_{i,j}$ may be necessary because TF $i$ varies little across the available data, and these large weights are not suggestive of a true regulatory interaction.

>>>>

\title{GRN inference in multiple species}
\author{Kari Y. Lam, Zachary M. Westrick, Christian Mueller, Lionel Christiaen, Rich Bonneau}
\maketitle

\begin{abstract}
Methods for learning gene regulatory networks are essential to our understanding of cell behavior and response to external stimuli. We are interested in improving the accuracy of our networks in order to do meaningful comparisons between networks. We present two approaches using fused regression, where we exploit available evolutionary information to constrain an undetermined system, and show that both result in more accurate networks. We show that fused L2 is most useful in the case where the networks are very similar (eg different cell types or closely related species), but where data is limited. We can leverage our knowledge about the relationship between genes to promote our understanding of the respective networks. When data is more readily available, adaptive fusion is useful for identifying conservation between networks.

\end{abstract}

@@

Where the second sum is over pairs of interactions with fusion constraints. Note that each term in the quantity being minimized is an L2 norm of a linear combination of gene expressions, so the entire problem can be solved through linear regression with a suitable design matrix. The regulators of a pair of genes can be solved independently as long as the pair is not linked by fusion constraints, or a chain of fusion constraints; thus the problem of cross-species GRN inference decomposes into a large number of manageable regression problems as long as the orthology mappings are sparse.

@@

For each fused $B^{S_1}_{g,k} \approx B^{S_2}_{h,l}$ we define:

\begin{equation} 
\theta(0)=0
\end{equation}
\begin{equation}
\theta(t) = \vert B^{S_1}_{g,k} - B^{S_2}_{h,l} \vert
\end{equation}

and introduce a fusion constraint $\lambda = \frac{p'(\theta(t))}{2\theta(t)} $

@@

\subsection{Beta scaling}
\\
In previous work, betas were rescaled as to form a matrix of confidence scores $S$ as follows
\begin{equation}
S_{i,j} = \frac{\sigma^2_{\text{full model for }y_j}}{\sigma^2_{\text{full model for }y_j \text{ without predictor }i}}
\end{equation}
Computing residuals with respect to the data alone would disregard information gained through fusion. Instead, we used an approximation
\begin{equation}
S_{i,j} = \frac{\sigma^2_{\text{full model for }y_j}}{\sigma^2_{\text{full model for }y_j} + B_{i,j}^2 \times var(TF_j)}
\end{equation}
When the residual is uncorrelated with the prediction, then removing regressor $i$ with weight $B_{i,j}^2$ from the model of $j$ will increase the residual by $B_{i,j}^2$ times the variance of regressor $i$. In the case of regularized regression, the regressor and residual need not be uncorrelated, so the approximation will not hold exactly. This scaling is similar to rescaling according to variance explained relative to an augmented design matrix that includes fusion constraints. We use an approximation because it is simpler to compute. The intuition is that a large $B_{i,j}$ may be necessary because TF $i$ varies little across the available data, and these large weights are not suggestive of a true regulatory interaction.
